---
title: "P&P analysis"
author: "Megumi Terada"
date: '2023-06-26'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(tidyverse)
library(rPraat)
library(tidyverse)

# articulation rate of each IPU of each interlocutor
load("D:/P&P/IPU-ArticulationRate.RData")
ipu <- int 

# mean articulation rate of the entire turn plus the articulation rate of the interlocutor's previous turn
load("D:/P&P/turns-ArticulationRate.RData")
turns <- int 

rm(int)

### exclude double data
all <- distinct(ipu %>% 
                  select(file, dyad, speaker, confederate, 
                         artRate, ipuOnset, ipuDur) %>% 
                  rename(type = speaker, 
                         interlocutor = confederate)) 

### add new columns

# column for participants
all$participantId <- sapply(strsplit(all$file,"_"), 
                            function(x) x[1])
all$participantId <- substr(all$participantId, 1, 3)

# column for confederates
all$confederateId <- sapply(strsplit(sapply(strsplit(all$file,"_"),
                                           function(x) x[1]), 
                               split = "p.."),
                                           function(x) x[2])
# additional information
all$language <- sapply(strsplit(all$file,"_"), function(x) x[2])
all$language <- substr(all$interlocutor, 1, 2)
all$proficiency <- substr(all$interlocutor, 3, 4)
all$proficiency[all$confederateId %in% c("c01","c03","c05","c07")] <- "L1"
all$round <- "1"
all$round[all$confederateId %in% c("c03","c04")] <- "2"
all$round[all$confederateId %in% c("c05","c06")] <- "3"
all$round[all$confederateId %in% c("c07","c08")] <- "4"

all$roundParticipant <- as.factor(all$participantId)
levels(all$roundParticipant) <- c(paste0("speaker ", rep(1:5,4)))

table(all$type, all$language)
```

```{r}
### check the data for each speaker

# distribution: participants when speaking to L1 and L2 confederates
ggplot(all %>% filter(type == "participant"), 
       aes(x = artRate, fill = language)) +
  geom_density(position = "identity", alpha = .5) +
  facet_grid(.~round)

# distribution: confederates when speaking to participants
ggplot(all %>% filter(type == "confederate"), 
       aes(x = artRate, fill = language)) +
  geom_density(position = "identity", alpha = .5) +
  facet_grid(.~round)

# boxplot: participants when speaking to L1 and L2 confederates
ggplot(all%>% filter(type == "participant"), 
       aes(y = artRate, x = language, fill = language)) +
  geom_boxplot() +
  facet_grid(.~round) +
  ylim(0, 12.5)

# boxplot: confederates when speaking to L1 and L2 confederates
ggplot(all%>% filter(type == "confederate"), 
       aes(y = artRate, x = language, fill = language)) +
  geom_boxplot() +
  facet_grid(.~round) +
  ylim(0, 12.5)
```

# Questions to the data
why duplicated?
what are the NA values?
outliers? (extreme high and low rate) - to be checked

# done
individual comparison 
- point plots indicating the intra-individual differences
- the corresponding speech rate of the L1 and L2 confederates

```{r}
# mean for participants
rate.mean <- all  %>%
  group_by(dyad, round, type, language, participantId, roundParticipant) %>% 
  summarise(mean.rate = mean(artRate, na.rm = TRUE))
  
# mean for confederates
rate.conf <- all  %>%
  group_by(round, language, confederateId) %>% 
  summarise(mean.rate = mean(artRate, na.rm = TRUE))

# point plot
ggplot() +
  geom_point(data = rate.mean %>% filter(type == "participant"), 
       aes(x = language, y = mean.rate, shape = roundParticipant), size = 2.5) +
  geom_path(data = rate.mean %>% filter(type == "participant"), 
            aes(x = language, y = mean.rate, group = roundParticipant)) +
  geom_point(data = rate.conf, 
       aes(x = language, y = mean.rate), size = 3, color = "red") +
  # geom_path(aes(group = speaker)) +
  facet_grid(.~ round) +
  ylim(3.5, 6.5) +
  labs(title = "Mean speech rate of each speaker", 
       x = "When speaking with confederates with", 
       y = "Mean speech rate (syllable per second)", 
       shape = "Speakers in each \nround (n=20)")
ggsave("NNAR_meanAR.jpg")

```

# first results
L1 and L2 confederates speak almost at the same rate -> no clear observation if the participants converge to L1/L2 speakers (at least in the first and second round)

in the third and forth rounds, the L2 confs speak more slowly resulting in convergence/foreigner talk, but not to differentiate

# another plot to be done
local articulation rate, difference at the beginning and the end?

```{r}
# change throughout the session: nope
ggplot(data = all %>% filter(type == "participant")) +
  geom_point(aes(x = ipuOnset, y = artRate, color = language)) +
  geom_smooth(aes(x = ipuOnset, y = artRate, color = language, group = language)) +
  facet_wrap(.~round)
```

