---
title: "Analysis CoNNAR - f0 and articulation rate"
author: "Tom Offrede"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(lmerTest)
library(broom)

folder <- paste0(here::here(), "/data/")
load(paste0(folder, "speechData.RData"))
dat <- dat |> 
  group_by()
```

# Notes to understand the analysis

* `confederate` refers to each individual confederate (B1, B2, C1-a, C1-b)
* `confederate type` refers to whether a confederate is an L1 or L2 speaker

* **Synchrony** refers to how a speaker changes their speech in accordance with the speech of the interlocutor: e.g. both produce higher f0, or both produce lower f0 -- here there is no approximation between the values of the interlocutors. This is translated into our models below as `speaker's feature ~ interlocutor's feature`. When evaluating if synchrony is affected by the confederate's proficiency, we add an interaction: `speaker's feature ~ interlocutor's feature : confederate's proficiency`. (Proficiency here can refer to either `confederate type`, i.e. L1 vs L2, or `accent rating`.)

* **Convergence** refers to increasing similarity between the two speakers' values over time. In our models, it's translated as `difference in features ~ turn`. Here, `turn` indicates the passage of time in the conversation. We measure influence of confederate's proficiency in the same way as for synchrony (with an interaction).

* **Register adaptation** below refers to the participants changing their articulation rate depending on who they're talking to: `speaker's feature ~ confederate's proficiency`.

# Summary of results

## Articulation rate

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
# took this code from https://stackoverflow.com/questions/19997242/simple-manual-rmarkdown-tables-that-look-good-in-html-pdf-and-docx
tabl <- "
| Type of adapatation | Did it occur overall? | Confederate effect?                                       |
|---------------------|:---------------------:|:---------------------------------------------------------:|
| Synchrony           | no                    | no                                                        |
| Convergence         | no                    | They converge more the lower the confederate's accent rating |
| Register            | no                    | no            |
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

## f0 mean    

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Type of adapatation | Did it occur overall? | Confederate effect? |
|---------------------|:---------------------:|:-------------------:|
| Synchrony           | yes                   | no                  |
| Convergence         | yes                 | slightly stronger convergence to **L2**          |
"
cat(tabl)
```

## f0 median    

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Type of adapatation | Did it occur overall? | Confederate effect?                                                     |
|---------------------|:---------------------:|:-----------------------------------------------------------------------:|
| Synchrony           | yes                   | no |
| Convergence         | yes            | no          |
"
cat(tabl)
```

## f0 SD    

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Type of adapatation | Did it occur overall? | Confederate effect?                     |
|---------------------|:---------------------:|:---------------------------------------:|
| Synchrony           | no                    | no                                      |
| Convergence         | marginally                    | no |
"
cat(tabl)
```

## f0 IQR    

```{r echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| Type of adapatation | Did it occur overall? | Confederate effect? |
|---------------------|:---------------------:|:-------------------:|
| Synchrony           | no                    | no                  |
| Convergence         | yes                   | no                  |
"
cat(tabl)
```
# Articulation rate

## Difference between confederates

### Confederate type

Most L2 confederates spoke slower than L1 confederates. Only one of the C1 confederates did not speak slower than L1 speakers, and the other C1 confederate's difference was also smaller. The B2 confederate spoke the slowest.

```{r}
ggplot(dat |> filter(role=="confederate"), aes(confederateType, turnAR))+
  geom_boxplot()

dat$confederate <- relevel(dat$confederate, ref="L1")
summary(lmer(turnAR ~ confederate + (1|file), dat |> filter(role=="confederate")))
summary(lmer(turnAR ~ confederateType + (1|file), dat |> filter(role=="confederate")))
```

### Accent ratings

The higher the accent rating, the slower the speech.

```{r}
ggplot(dat |> filter(role=="confederate"), aes(as.factor(accentRating), turnAR))+
  geom_boxplot()

summary(lmer(turnAR ~ accentRatingc + (1|file), dat |> filter(role=="confederate")))
```

## Synchrony

The model below shows that the participants don't synchronize  to the confederate.

```{r}
summary(m0 <- lmer(turnAR ~ prevARc + (1|file), dat |> filter(role=="participant")))
```

### Confederate effect

And they don't synchronize regardless of who the confederate is.
Considering accent rating instead of `confederate type` also didn't bring out any effects.

```{r}
summary(m1 <- lmer(turnAR ~ prevARc : confederateType + (1|file), dat |> filter(role=="participant")))
summary(m2 <- lmer(turnAR ~ prevARc : accentRatingc + (1|file), dat |> filter(role=="participant")))

anova(m0, m1)
anova(m0, m2)
```

## Convergence

Participants don't converge on articulation rate to the confederates (i.e., the difference between their features doesn't change across time).

```{r}
summary(m0 <- lmer(artRateDiff ~ turn + (1|file), dat |> filter(role=="participant")))
```

### Confederate effect

`Confederate type` didn't improve the model, but `accent rating` did! Participants converge on articulation rate more the lower the confederate's accent rating.

```{r}
summary(m1 <- lmer(artRateDiff ~ turn : confederateType + (1|file), dat |> filter(role=="participant")))
summary(m2 <- lmer(artRateDiff ~ turn : accentRatingc + (1|file), dat |> filter(role=="participant")))

anova(m0, m1)
anova(m0, m2)
```


## Register adaptation

Participants did not speak differently to L1 vs L2 speakers.

```{r}
summary(lmer(turnAR ~ confederateType + (1|file), dat |> filter(role=="participant")))
summary(lmer(turnAR ~ accentRatingc + (1|file), dat |> filter(role=="participant")))
```


# f0 mean

## Synchrony

Participants seem to synchronize on f0 mean to the confederate.

```{r}
summary(m0 <- lmer(f0meanz ~ prevf0meanC + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

The model without `confederate type` is better than the one with, and there is no effect with `accent rating` So participants do **not** synchronize differently to L1 vs L2 confederates.

```{r}
summary(m1 <- lmer(f0meanz ~ prevf0meanC : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0meanz ~ prevf0meanC : accentRatingc + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

## Convergence

Participants also converge to the confederates.

```{r}
summary(m0 <- lmer(f0meanDiff ~ turn + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

`Confederate type` seems to improve the model (lower AIC), but the direction of the effect is the same. It seems that participants converge more to L2 than to L1 speakers though.

```{r}
summary(m1 <- lmer(f0meanDiff ~ turn : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0meanDiff ~ turn : accentRating + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

# f0 median

## Synchrony

As for f0 mean: participants synchronize to their interlocutors.

```{r}
summary(m0 <- lmer(f0med ~ prevf0medC + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

`Confederate type` did not improve the model. Participants synchronized on f0 median to the same extent.

```{r}
summary(m1 <- lmer(f0med ~ prevf0medC : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0med ~ prevf0medC : accentRatingc + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

## Convergence

Participants converged on f0 median.

```{r}
summary(m0 <- lmer(f0medDiff ~ turn + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

`Confederate type` made the model worse.

```{r}
summary(m1 <- lmer(f0medDiff ~ turn : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0medDiff ~ turn : accentRating + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

# f0 SD

## Synchrony

Participants did not synchronize on f0 SD.

```{r}
summary(m0 <- lmer(f0sd ~ prevf0sdC + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

Adding `confederate` to the model didn't improve it enough. There was no robust by-confederate effect.

```{r}
summary(m1 <- lmer(f0sd ~ prevf0sdC : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0sd ~ prevf0sdC : accentRatingc + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

## Convergence

Participants converged slightly on f0 SD.

```{r}
summary(m0 <- lmer(f0sdDiff ~ turn + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

Including `confederate type` in the model increased AIC. There was no robust confederate effect.

```{r}
summary(m1 <- lmer(f0sdDiff ~ turn : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0sdDiff ~ turn : accentRating + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

# f0 IQR

## Synchrony

Participants did not synchronize on f0 IQR.

```{r}
summary(m0 <- lmer(f0IQR ~ prevf0IQRC + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

There was no by-confederate effect either.

```{r}
summary(m1 <- lmer(f0IQR ~ prevf0IQRC : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0IQR ~ prevf0IQRC : accentRatingc + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```

## Convergence

Participants converged on f0 IQR.

```{r}
summary(m0 <- lmer(f0IQRDiff ~ turn + (1|file), dat |> filter(role == "participant")))
```

### Confederate effect

Including `confederate type` and `accent rating` in the model increased AIC (i.e. made the model worse). No confederate effect.

```{r}
summary(m1 <- lmer(f0IQRDiff ~ turn : confederateType + (1|file), dat |> filter(role == "participant")))
summary(m2 <- lmer(f0IQRDiff ~ turn : accentRating + (1|file), dat |> filter(role == "participant")))

anova(m0, m1)
anova(m0, m2)
```


# Some individual effects

```{r}
# # f0 median
# 
# folderInd <- paste0(here::here(), "/individual/f0median/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(f0medDiff ~ turn : confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(turn, f0medDiff))+
#     geom_boxplot()+
#     geom_point()+
#     facet_wrap(~confederateType)+
#     geom_smooth(method="lm")+
#     ggtitle(paste(s, "; L1:", format(round(m$statistic[2], 3)), "; L2:", format(round(m$statistic[3], 3))))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
# 
# # f0 mean
# 
# folderInd <- paste0(here::here(), "/individual/f0mean/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(f0meanDiff ~ turn : confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(turn, f0meanDiff))+
#     geom_boxplot()+
#     geom_point()+
#     facet_wrap(~confederateType)+
#     geom_smooth(method="lm")+
#     ggtitle(paste(s, "; L1:", format(round(m$statistic[2], 3)), "; L2:", format(round(m$statistic[3], 3))))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
# 
# # f0 SD
# 
# folderInd <- paste0(here::here(), "/individual/f0SD/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(f0sdDiff ~ turn : confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(turn, f0sdDiff))+
#     geom_boxplot()+
#     geom_point()+
#     facet_wrap(~confederateType)+
#     geom_smooth(method="lm")+
#     ggtitle(paste(s, "; L1:", format(round(m$statistic[2], 3)), "; L2:", format(round(m$statistic[3], 3))))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
# 
# # f0 IQR
# 
# folderInd <- paste0(here::here(), "/individual/f0IQR/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(f0IQRDiff ~ turn : confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(turn, f0IQRDiff))+
#     geom_boxplot()+
#     geom_point()+
#     facet_wrap(~confederateType)+
#     geom_smooth(method="lm")+
#     ggtitle(paste(s, "; L1:", format(round(m$statistic[2], 3)), "; L2:", format(round(m$statistic[3], 3))))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
```

```{r}
# # articulation rate - register adaptation
# 
# folderInd <- paste0(here::here(), "/individual/register-ArticRate/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(turnAR ~ confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(confederateType, turnAR))+
#     geom_boxplot()+
#     ggtitle(paste(s, "t = ", format(round(m$statistic[2], 3)), "(Confederate:", dat$confederate[dat$speaker==s & !grepl("L1", dat$confederate)], ")"))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
# 
# # articulation rate - convergence
# 
# folderInd <- paste0(here::here(), "/individual/conv-ArticRate/")
# for(s in unique(dat$speaker[dat$role=="participant"])){
#   m <- tidy(lm(artRateDiff ~ turn : confederateType, dat |> filter(speaker==s)))
#   ggplot(dat |> filter(speaker==s), aes(turn, artRateDiff))+
#     geom_boxplot()+
#     geom_point()+
#     geom_smooth(method="lm")+
#     facet_wrap(~confederateType)+
#     ggtitle(paste(s, "; L1:", format(round(m$statistic[2], 3)), "; L2:", format(round(m$statistic[3], 3)), "(Conf:", dat$confederate[dat$speaker==s & !grepl("L1", dat$confederate)], ")"))
#   ggsave(filename = paste0(folderInd, s, ".png"))
# }
```


# Figures for poster

```{r}
ggplot(dat |> filter(role=="participant"), aes(turn, f0IQRDiff))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~confederateType, scales = "free")+
  labs(x = "Time", y="Difference in f0 IQR", title="Convergence on f0 IQR")+
  theme_bw()+
  theme(plot.title = element_text(hjust=0.5),
        )
ggsave(filename=paste0(here::here(), "/figures/f0-conv.png"))
```

# Session Info

```{r}
sessionInfo()
```

