---
title: "Preprocessing"
author: "Tom Offrede"
date: "2023-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read TextGrids, which contain annotation of turn duration and phonetic transcription, and create dataset with speech rate in each audio file.
In addition, read txt files containing f0 information per turn, and include that in the dataset.

```{r}
library(stringr)
library(tidyverse)
library(rPraat)

folderGen <- paste0(here::here(), "/data/")
# folderData <- paste0(here::here(), "/data/updated2/")
folderData <- paste0("C:/Users/offredet/Documents/1HU/CoNNAR/CoNNAR/data/all/")
# folderTurn <- paste0(here::here(), "/data/final/")

filesTG <- list.files(folderData, "TextGrid")
filesTXT <- list.files(folderData, "\\.txt")

ff <- data.frame(filesTG, filesTXT) |>
  mutate(worked = ifelse(gsub(".TextGrid", "", filesTG) == gsub(".txt", "", filesTXT), # to make sure the files are matching
                         "worked!", "NO!!!!!"))
table(ff$worked)
```

Get dataset with IPUs, turns, articulation rate, and f0

```{r}
int <- data.frame(matrix(nrow=0, ncol=16))
names(int) <- c("file", "dyad", "speaker", "confederate", "turn", "intervalIPU", "label", "ipuOnset", "ipuOffset", "ipuDur", "deleteDur", "vowels", "f0mean", "f0med", "f0IQR", "f0sd")

for(f in 1:nrow(ff)){
  tg <- tg.read(paste0(folderData, ff$filesTG[f]), encoding = detectEncoding(paste0(folderData, ff$filesTG[f])))
  
  for(i in 1:tg.getNumberOfIntervals(tg, "IPU")){
    if(tg.getLabel(tg, "IPU", i) == "ipu"){
      del <- 0
      vowels <- 0
      startIPU <- tg.getIntervalStartTime(tg, "IPU", i)
      endIPU <- tg.getIntervalEndTime(tg, "IPU", i)
      
      int[nrow(int)+1,] <- c(gsub(".TextGrid", "", ff$filesTG[f]),
                             str_split(ff$filesTG[f], fixed("_"))[[1]][[1]],
                             ifelse(grepl("_ch1", ff$filesTG[f]), "confederate", ifelse(grepl("_ch2", ff$filesTG[f]), "participant", "error")),
                             str_split(ff$filesTG[f], fixed("_"))[[1]][[2]],
                             NA, # this will be turn
                             i,
                             tg.getLabel(tg, "IPU", i),
                             startIPU,
                             endIPU,
                             tg.getIntervalDuration(tg, "IPU", i),
                             NA, NA, NA, NA, NA, NA)
      
      # save turns
      for(t in 1:tg.getNumberOfIntervals(tg, "turns")){
        if(tg.getIntervalStartTime(tg, "turns", t) <= tg.getIntervalStartTime(tg, "IPU", i)){
          if(tg.getIntervalEndTime(tg, "turns", t) >= tg.getIntervalEndTime(tg, "IPU", i)){
            int$turn[nrow(int)] <- tg.getLabel(tg, "turns", t)
          }
        }
      }
      
      # save the periods of time to delete due to them being annotated as <usb>, <nib>, <pseudo>, or <p:>
      for(p in c(tg.getIntervalIndexAtTime(tg, "MAU", startIPU):tg.getIntervalIndexAtTime(tg, "MAU", (endIPU-0.00001)))){
        if(grepl("<", tg.getLabel(tg, "MAU", p))){
          del <- as.numeric(del + as.numeric(tg.getIntervalDuration(tg, "MAU", p))) # if there are multiple "<>" intervals, we add up all of them
        }
        
        if(!grepl("<", tg.getLabel(tg, "MAU", p))){
          if(grepl("@|2|9|a|E|e|I|i|O|o|U|u|y|Y", tg.getLabel(tg, "MAU", p))){
            vowels <- vowels + 1
          }
          if(grepl("6", tg.getLabel(tg, "MAU", p))){ # if the vowel is "6", it should only be counted as a different syllable if the previous phone wasn't a vowel
            if(!grepl("@|2|9|a|E|e|I|i|O|o|U|u|y|Y", tg.getLabel(tg, "MAU", p-1))){
              vowels <- vowels + 1
            }
          }
        }
      }
      
      int$deleteDur[nrow(int)] <- as.numeric(del)
      int$vowels[nrow(int)] <- as.numeric(vowels)
    }
  }
}

int <- int %>% # for some reason the `|>` pipe is giving an error in this chunk, so using `%>%`
  filter(turn != "") %>% 
  group_by(file) %>% 
  mutate(IPU = 1:n()) %>% 
  ungroup %>% 
  mutate_at(c("ipuDur", "deleteDur", "vowels", "ipuOnset", "ipuOffset"), as.numeric) %>% 
  mutate(realDur = ipuDur - deleteDur,
         artRate = ifelse(realDur != 0 & vowels != 0, vowels / realDur, NA),
         turn = gsub("turn", "", turn)) %>% 
  group_by(file, turn) %>% 
  mutate(turnAR = mean(artRate, na.rm=TRUE),
         turnOnset = as.numeric(min(ipuOnset)),
         turnOffset = as.numeric(max(ipuOffset))) %>% 
  ungroup()

for(f in 1:nrow(ff)){
  txt <- read.table(paste0(folderData, ff$filesTXT[f]), header = TRUE, na.strings = "--undefined--") |> 
    mutate(meanZ = (f0mean - mean(f0mean, na.rm=TRUE)) / sd(f0mean, na.rm=TRUE),
           medZ = (f0med - mean(f0med, na.rm=TRUE)) / sd(f0med, na.rm=TRUE),
           sdZ = (f0sd - mean(f0sd, na.rm=TRUE)) / sd(f0sd, na.rm=TRUE),
           iqrZ = (f0IQR - mean(f0IQR, na.rm=TRUE)) / sd(f0IQR, na.rm=TRUE)) |> 
    mutate(f0mean = ifelse(abs(meanZ) > 2.5, NA, f0mean),
           f0med = ifelse(abs(medZ) > 2.5, NA, f0med),
           f0sd = ifelse(abs(sdZ) > 2.5, NA, f0sd),
           f0IQR = ifelse(abs(iqrZ) > 2.5, NA, f0IQR))
  for(i in 1:nrow(int)){
    if(int$file[i] == gsub(".txt", "", ff$filesTXT[f])){
      for(r in 1:nrow(txt)){
        if(abs(txt$onset[r] - int$turnOnset[i]) < 0.01){
          if(abs(txt$offset[r] - int$turnOffset[i]) < 0.01){
            int$f0mean[i] <- as.numeric(txt$f0mean[r])
            int$f0med[i] <- as.numeric(txt$f0med)[r]
            int$f0IQR[i] <- as.numeric(txt$f0IQR)[r]
            int$f0sd[i] <- as.numeric(txt$f0sd)[r]
          }
        }
      }
    }
  }
}

# save(int, file=paste0(folderGen, "IPU-ArticulationRate.RData"))
SAVE <- int
```

Add confederates' accent ratings

```{r}
int <- int |> 
  mutate(accentRating = case_when(
    grepl("c01|c03|c05|c07", file) ~ 0,
    grepl("c02", file) ~ 2.35,
    grepl("c04", file) ~ 2.57,
    grepl("c06", file) ~ 5.44,
    grepl("c08", file) ~ 3.26
  ),
  confederate = case_when(
    grepl("c02", file) ~ "C1-a",
    grepl("c04", file) ~ "C1-b",
    .default = as.character(confederate)
  ))
```

Rename and adjuts variables

```{r}
int <- int |> 
  rename(role = speaker) |> 
  mutate(speaker = ifelse(role=="participant", substr(file, 1, 3), substr(file, 4, 6)),
         interlocutor = ifelse(role=="confederate", substr(file, 1, 3), substr(file, 4, 6))) |> 
  mutate_at(c("turn", "f0mean", "f0med", "f0IQR", "f0sd"), as.numeric) |>
  group_by(speaker) |> 
  mutate(f0meanz = (f0mean - mean(f0mean, na.rm=TRUE)) / sd(f0mean, na.rm=TRUE), # z score f0 measures because the dyads were often mixed-sex
         f0medz = (f0med - mean(f0med, na.rm=TRUE)) / sd(f0med, na.rm=TRUE),
         f0sdz = (f0sd - mean(f0sd, na.rm=TRUE)) / sd(f0sd, na.rm=TRUE),
         f0IQRz = (f0IQR - mean(f0IQR, na.rm=TRUE)) / sd(f0IQR, na.rm=TRUE)
  )
```

Annotate the articulation rate and f0 of interlocutor's previous turn in the dataset

```{r}
dat <- int |> 
  group_by(file, turn) |> 
  filter(!duplicated(turn)) |> 
  ungroup() |> 
  select(-c(intervalIPU, label, ipuOnset, ipuOffset, ipuDur, deleteDur, IPU, artRate)) |> 
  mutate(firstSp = NA)

# determine who is the first speaker in each dyad
for(f in unique(dat$dyad)){
  dat$firstSp[dat$dyad==f] <- unique(as.character(dat$speaker[dat$dyad==f & dat$turnOnset == min(dat$turnOnset[dat$dyad==f])]))
}

dat <- dat |> 
  group_by(file) |> 
  mutate_at(c("turn", "f0meanz", "f0medz", "f0IQRz", "f0sdz"), as.numeric) |> 
  mutate(prevTurn = ifelse(speaker == firstSp, turn-1, turn)) |>
  mutate(prevTurn = ifelse(speaker == firstSp & turn == 1, NA, prevTurn),
         prevAR = NA, prevf0mean = NA, prevf0med = NA, prevf0IQR = NA, prevf0sd = NA) |>
  ungroup()

for(i in 1:nrow(dat)){
  prevAR <- as.numeric(dat$turnAR[dat$dyad == dat$dyad[i] &
                                    dat$speaker != dat$speaker[i] &
                                    dat$turn == dat$prevTurn[i]])
  prevf0mean <- as.numeric(dat$f0meanz[dat$dyad == dat$dyad[i] &
                                         dat$speaker != dat$speaker[i] &
                                         dat$turn == dat$prevTurn[i]])
  prevf0med <- as.numeric(dat$f0medz[dat$dyad == dat$dyad[i] &
                                       dat$speaker != dat$speaker[i] &
                                       dat$turn == dat$prevTurn[i]])
  prevf0IQR <- as.numeric(dat$f0IQRz[dat$dyad == dat$dyad[i] &
                                       dat$speaker != dat$speaker[i] &
                                       dat$turn == dat$prevTurn[i]])
  prevf0sd <- as.numeric(dat$f0sdz[dat$dyad == dat$dyad[i] &
                                     dat$speaker != dat$speaker[i] &
                                     dat$turn == dat$prevTurn[i]])
  
  if(!purrr::is_empty(prevAR)){
    if(!any(is.na(prevAR))){
      dat$prevAR[i] <- as.numeric(prevAR)
    }
  }
  if(!purrr::is_empty(prevf0mean)){
    if(!any(is.na(prevf0mean))){
      dat$prevf0mean[i] <- as.numeric(prevf0mean)
    }
  }
  if(!purrr::is_empty(prevf0med)){
    if(!any(is.na(prevf0med))){
      dat$prevf0med[i] <- as.numeric(prevf0med)
    }
  }
  if(!purrr::is_empty(prevf0IQR)){
    if(!any(is.na(prevf0IQR))){
      dat$prevf0IQR[i] <- as.numeric(prevf0IQR)
    }
  }
  if(!purrr::is_empty(prevf0sd)){
    if(!any(is.na(prevf0sd))){
      dat$prevf0sd[i] <- as.numeric(prevf0sd)
    }
  }
}

# save(int, file=paste0(folderGen, "turns-ArticulationRate.RData"))
```

Calculate difference between the interlocutors' values

```{r}
dat <- dat |> 
  mutate(artRateDiff = abs(turnAR - prevAR),
         f0meanDiff = abs(f0meanz - prevf0mean),
         f0medDiff = abs(f0medz - prevf0med),
         f0IQRDiff = abs(f0IQRz - prevf0IQR),
         f0sdDiff = abs(f0sdz - prevf0sd))
```

Fix format of variables (factor, numeric etc)

```{r}
dat <- dat |> 
  mutate_at(c("file", "dyad", "speaker", "confederate"), as.factor)
```

Center predictors

```{r}
dat <- dat |> 
  mutate(prevf0meanC = prevf0mean - mean(prevf0mean, na.rm=TRUE),
         prevf0medC = prevf0med - mean(prevf0med, na.rm=TRUE),
         prevf0IQRC = prevf0IQR - mean(prevf0IQR, na.rm=TRUE),
         prevf0sdC = prevf0sd - mean(prevf0sd, na.rm=TRUE),
         prevARc = prevAR - mean(prevAR, na.rm=TRUE),
         accentRatingc = accentRating - mean(accentRating, na.rm=TRUE),
         confederateType = ifelse(grepl("L1", confederate), "L1", "L2")) # create a variable `confederateType`
```


Save dataset

```{r}
save(dat, file=paste0(folderGen, "speechData.RData"))
```


Plot the differences

```{r}
# articulation rate

ggplot(dat, aes(artRateDiff, turn))+
  geom_point()+
  geom_smooth(method="lm")+ # to see non-linear effect, use method="loess"
  facet_wrap(~dyad)

# f0 mean

ggplot(dat, aes(f0meanDiff, turn))+
  geom_point()+
  geom_smooth(method="lm")+ # to see non-linear effect, use method="loess"
  facet_wrap(~dyad)

# f0 med

ggplot(dat, aes(f0medDiff, turn))+
  geom_point()+
  geom_smooth(method="lm")+ # to see non-linear effect, use method="loess"
  facet_wrap(~dyad)

# f0 max

ggplot(dat, aes(f0IQRDiff, turn))+
  geom_point()+
  geom_smooth(method="lm")+ # to see non-linear effect, use method="loess"
  facet_wrap(~dyad)

# f0 sd

ggplot(dat, aes(f0sdDiff, turn))+
  geom_point()+
  geom_smooth(method="lm")+ # to see non-linear effect, use method="loess"
  facet_wrap(~dyad)
```


```{r}
# to check file naming

# names <- data.frame(matrix(nrow=0, ncol=4))
# names(names) <- c("1", "2", "3", "4")
# for(f in files){
#   s <- str_split(gsub(".TextGrid", "", f), fixed("_"))[[1]]
#   names[nrow(names)+1,] <- c(s[[1]], s[[2]], s[[3]], s[[4]])
# }
# names
```

```{r}
# to check what phones there are and what words they come from

# phones <- data.frame(matrix(nrow=0, ncol=4))
# names(phones) <- c("file", "phone", "word", "wordPhon")
# for(f in files){
#   tg <- tg.read(paste0(folderData, f), encoding = detectEncoding(paste0(folderData, f)))
#   
#   for(i in 1:tg.getNumberOfIntervals(tg, "MAU")){
#     if(tg.getLabel(tg, "MAU", i) != ""){
#       phones[nrow(phones)+1,] <- c(f,
#                                    tg.getLabel(tg, "MAU", i),
#                                    tg.getLabel(tg, "ORT-MAU", tg.getIntervalIndexAtTime(tg, "ORT-MAU", as.numeric(tg.getIntervalStartTime(tg, "MAU", i))+0.000001)),
#                                    tg.getLabel(tg, "KAN-MAU", tg.getIntervalIndexAtTime(tg, "KAN-MAU", as.numeric(tg.getIntervalStartTime(tg, "MAU", i))+0.000001)))
#     }
#   }
# }
# phones <- phones |> filter(phone!="<p:>")
# p <- aggregate(word ~ phone, phones, paste, collapse=",")
```